A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-02-22 jacobian_calc_CMAME.tex
supported the assertion of byte-copyability being significant with evidence that should matter to stampede users

The CS method was shown to be highly accurate in calculating tangent-stiffness
matrices by the standards of this study, and it is slower than other methods
for a naive implementation. In the context of future use of CS in
\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
while AD requires a complicated serialization and deserialization in order to
function as byte-copyable. This is fine for copying in between MPI processes,
but deserialization on an accelerator or GPU is very difficult, as this would
require compiling a version of the AD library being used that were compatible
with the accelerator or GPU being used.  What this means is
that if \emph{Peridigm} or any simulation software were to be reformulated to
take advantage of accelerators or GPUs, the flexibility of complex-step owed to
its simple implementation may make it more viable for that application given
its high accuracy. Additional performance gains may be achieved through the use
of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
use of complex data types natively, and alleviate the expense of the
dynamic-casts that were utilized in this study.  As a final comment, it should
be noted that from a code development perspective the CS method may be the
easiest of the methods to implement; It only requires one functional
evaluation, there is no heuristic choice of probe distance, and there is no
dependence on an external AD library.  From this perspective, the method may
still be an attractive method for TS matrix evaluation. \add[Michael]{However, at least for 
the models producing the results shown here, each of the
methods returned quadratic convergence, indicating each was sufficiently
accurate. 'Marching orders' for the analyst would be to select the swiftest
method, while it would seem accuracy above what is adequate does not enhance
the convergence rate or iteration speed until either classical Newton's convergence
conditions or the approximate case conditions are violated.}

>>>>

The CS method was shown to be highly accurate in calculating tangent-stiffness
matrices by the standards of this study, and it is slower than other methods
for a naive implementation. In the context of future use of CS in
\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
while AD requires a complicated serialization and deserialization in order to
function as byte-copyable. This is fine for copying in between MPI processes,
but deserialization on an accelerator or GPU is very difficult, as this would
require compiling a version of the AD library being used that were compatible
with the accelerator or GPU being used. A working definition of byte-copyable
is that the collection of data to transfer is made of identically sized pieces.
Byte-copyability is significant for example because it is a requirement for the
basic offload mode provided for Intel MIC accelerator boards which are
installed in \emph{Stampede}, where the alternative mode requires "more
extensive" effort \cite{intel_byte_copyable}. What this means is that if
\emph{Peridigm} or any simulation software were to be reformulated to take
advantage of accelerators or GPUs, the flexibility of complex-step owed to its
simple implementation may make it more viable for that application given its
high accuracy. Additional performance gains may be achieved through the use of
the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
use of complex data types natively, and alleviate the expense of the
dynamic-casts that were utilized in this study.  As a final comment, it should
be noted that from a code development perspective the CS method may be the
easiest of the methods to implement; It only requires one functional
evaluation, there is no heuristic choice of probe distance, and there is no
dependence on an external AD library.  From this perspective, the method may
still be an attractive method for TS matrix evaluation. \add[Michael]{However,
at least for the models producing the results shown here, each of the
methods returned quadratic convergence, indicating each was sufficiently
accurate. 'Marching orders' for the analyst would be to select the swiftest
method, while it would seem accuracy above what is adequate does not
enhance the convergence rate or iteration speed until either classical
Newton's convergence conditions or the approximate case conditions are
violated.}
